Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-08-16 draft.tex
Filled a few blanks, improved error rate description

\paragraph{The Error Term. }
Normally the error rate for a classification is calculated with $error = \frac{n_{classifiedCorrectly}} {n}$ where $n$ is the number of subjects.
%%
This metric says nothing about the discriminative power of each manifestation.
%%
If for example a variable indicating a disease is negative for 90\% of the subject and the classifier simply assigns all subjects to \emph{not ill} the error metric would yield an error of 10\%, even though it is very bad.
%%
Therefore the discriminative power of each dimension in the equation
%error = 1 - \frac{\[ \sum_{m=0}^M \] \frac{M_{classifiedCorrectly}}{M_all}}{number of dimensions}
\begin{equation}
error = 1 - \frac{\sum_{m=0}^M \frac{m_{classifiedCorrectly}}{m_{all}}}{number_{dimensions}}
\end{equation}
where $M$ represents each manifestation of a variable.
%%
The error is scaled to denote perfect classification with 0 and 1 is equal to random selection.
% The error rate has to be judged depending on the number of classes of the target variable.
% %%
% An error rate of 50\% for a target variable with two manifestations makes the classifier equal to throwing a coin.
% %%
% The same error rate is acceptable a target variable with more than two manifestations.
% %%
% Picking a value for a variable of four manifestations with an error rate of 50\% is better than the random selection (error rate of 75\%).
% %%
% The following scales the error rate so that 0 denotes perfect classification and 1 is equal to random selection.
% %%
% \begin{equation}
% error_{normalized} = \frac{error}{100 - \frac{100}{class number}}
% \end{equation}
%%
This allows for comparability of error rates between variables.

@@

Many follow-up variables are sparse, such as medication of diabetes or reason of early retirement.
%%
The classification algorithm may produce higher accuracy for variables with less subjects due to the small sample size.
%%
This makes these results less reliable.
%%
We map the number of subjects associated with a variable to point diameter in the scatter plot.
%%
This allows to instantly assess the reliability of the result.
%%
\com{ToDo: Manche Variablen rausgeschmissen, weil sie zu wenig Probanden enthielten, oder Auspraegungen nur duenn besetzt waren!}

\com{ToDo: SQRT Skalierung der Treesize-Dimension, da hauptsaechlich geringe Groessen von interesse sind, ausreisser aber diesen Bereich schwer durchschaubar machen}
% %%
%
% %%
% Abstracting the classification results into two continuos variables allows us to visualize the result of 134 classifications in one scatter plot.
% %%
% By mapping the group affiliation (in our case body-size) to color or shape, we can even plot all 402 results in a single visualization (Fig.~\ref{fig:decision-tree} a).
%%
\subsubsection{Dummy Variables}
Dummy variables abstract a categorical variable with multiple manifestations into several binary variables.
%%
Each binary variables encodes the presence of a manifestation.
%%
For example, a pain indicator variable ranging from \emph{1 - no pain} to \emph{4 - large pain} is subdivided into four binary variables (\emph{No pain - Yes/No} to \emph{Large Pain - Yes/no}).
%%
One subject can only have one of these variables set to true.
%%
This is useful for our classification, because it allows to determine which manifestation of a variable can be described best using the image data parameter.
%%\com{ToDo: Dummy-Variables}
%%------------
\subsubsection{Interaction With the Visualization}
%%
The described visualization provides a good overview over the classification results.
%%
We however still want to be able to look display \emph{details-on-demand} \ref{shneiderman1996} and examine a decision tree in detail.
%%
This is realized by clicking on a entry on the visualization, which then displays the corresponding decision tree in detail (Fig.~\ref{fig:decision-tree} b).
%%
This allows for sequentially analyzing the classifications.
%%------------
\subsubsection{Implementation}
All analysis are carried out using \texttt{R}, a popular programming language for statistical calculations and visualizations.
%%
The interactive visualizations are realized using the \texttt{ggvis}\footnote{GGVIS Footnote} package.
%%
As opposed to the standard \texttt{R} plots, \texttt{ggvis} allows to adjust visualization parameters using UI controls, such as sliders.
%%
We implemented our application as web based prototype using \texttt{R Shiny}\footnote{R Shiny footnote}.
%%
The package combines the power of \texttt{R} with advantages of web-based applications.
%%
It allows us to quickly exchange results with our collaborating epidemiological partners.
%%
They can try out the technique without installing anything.
%%
Exchanging software becomes as easy as exchanging a hyperlink.
%%---------
\subsection{Results}
We ran the analysis using different subject groups:
\begin{enumerate}
	%%
	\item All subjects,
	%%
	\item subdivision into \emph{male} and \emph{female},
	%%
	\item subdivision into \emph{age quantiles}, yielding the groups (21, 43.8] (43.8, 54] (54, 64] (64, 90],
	%%
	\item subdivision into \emph{size quantiles}, yielding the groups (139, 164], (164, 171], (171, 177], (177, 202].
	%%
\end{enumerate}
We plotted each group twice.
%%
The first plot shows all original variables.
%%
The second plot shows all categorical variables transformed into dichotomous dummy variables.

>>>>

\paragraph{The Error Term. }
Normally the error rate for a classification is calculated with $error = \frac{n_{classifiedCorrectly}} {n}$ where $n$ is the number of subjects.
%%
The metric usually works well for normally distributed variables.
%%
It distorts the result for other distribution types.
%%
If for example a variable indicating a disease is negative for 90\% of the subject and the classifier simply assigns all subjects to \emph{not ill} the error metric would yield an error of 10\%, even though it is very bad.
%%
Our error term therefore incorporates the discriminative power of each manifestation and is denoted as follows:
%%
%Therefore the discriminative power of each dimension in the equation
%error = 1 - \frac{\[ \sum_{m=0}^M \] \frac{M_{classifiedCorrectly}}{M_all}}{number of dimensions}
\begin{equation}
error = 1 - \frac{\sum_{m=0}^M \frac{m_{classifiedCorrectly}}{m_{all}}}{number_{dimensions}}
\end{equation}
$M$ represents the set of manifestation of each variable.
%%
The error is scaled to denote perfect classification with 0 and 1 is equal to random selection.
% The error rate has to be judged depending on the number of classes of the target variable.
% %%
% An error rate of 50\% for a target variable with two manifestations makes the classifier equal to throwing a coin.
% %%
% The same error rate is acceptable a target variable with more than two manifestations.
% %%
% Picking a value for a variable of four manifestations with an error rate of 50\% is better than the random selection (error rate of 75\%).
% %%
% The following scales the error rate so that 0 denotes perfect classification and 1 is equal to random selection.
% %%
% \begin{equation}
% error_{normalized} = \frac{error}{100 - \frac{100}{class number}}
% \end{equation}
%%
It allows for comparability of error rates between variables with different manifestation count.

@@

Many follow-up variables are sparse, such as medication of diabetes or reason of early retirement.
%%
The classification algorithm may produce higher accuracy for variables with less subjects due to the small sample size.
%%
This makes these results less reliable.
%%
We map the number of subjects associated with a variable to point diameter in the scatter plot.
%%
This allows to instantly assess the reliability of the result.
%%

We apply a square root scale for the tree size axis to highlight data points with few decision rules.
%%
Outlier results with very large decision trees distort otherwise the resulting plot.
%%

\com{Needs to be moved}
\add{We removed variables with less than 100 subjects, since they have to few subjects to be statistically significant.}
\com{ToDo: Manche Variablen rausgeschmissen, weil sie zu wenig Probanden enthielten, oder Auspraegungen nur duenn besetzt waren!}

%\com{ToDo: SQRT Skalierung der Treesize-Dimension, da hauptsaechlich geringe Groessen von interesse sind, ausreisser aber diesen Bereich schwer durchschaubar machen}
% %%
%
% %%
% Abstracting the classification results into two continuos variables allows us to visualize the result of 134 classifications in one scatter plot.
% %%
% By mapping the group affiliation (in our case body-size) to color or shape, we can even plot all 402 results in a single visualization (Fig.~\ref{fig:decision-tree} a).
%%
\subsubsection{Dummy Variables}
Dummy variables abstract a categorical variable with multiple manifestations into several binary variables.
%%
Each binary variables encodes the presence of a manifestation.
%%
For example, a pain indicator variable ranging from \emph{1 - no pain} to \emph{4 - large pain} is subdivided into four binary variables (\emph{No pain - Yes/No} to \emph{Large Pain - Yes/no}).
%%
One subject can only have one of these variables set to true.
%%
This is useful for our classification, because it allows to determine which manifestation of a variable can be described best using the image data parameter.
%%\com{ToDo: Dummy-Variables}
%%------------
\subsubsection{Interaction With the Visualization}
%%
The described visualization provides a good overview over the classification results.
%%
We however still want to be able to look display \emph{details-on-demand} \ref{shneiderman1996} and examine a decision tree in detail.
%%
This is realized by clicking on a entry on the visualization, which then displays the corresponding decision tree in detail (Fig.~\ref{fig:decision-tree} b).
%%
This allows for sequentially analyzing the classifications.
%%------------
\subsubsection{Implementation}
All analysis are carried out using \texttt{R}, a popular programming language for statistical calculations and visualizations.
%%
The interactive visualizations are realized using the \texttt{ggvis}\footnote{GGVIS Footnote} package.
%%
As opposed to the standard \texttt{R} plots, \texttt{ggvis} allows to adjust visualization parameters using UI controls, such as sliders.
%%
We implemented our application as web based prototype using \texttt{R Shiny}\footnote{R Shiny footnote}.
%%
The package combines the power of \texttt{R} with advantages of web-based applications.
%%
It allows us to quickly exchange results with our collaborating epidemiological partners.
%%
They can try out the technique without installing anything.
%%
Exchanging software becomes as easy as exchanging a hyperlink.
%%---------
\subsection{Results}
We ran the analysis using different subject groups:
\begin{enumerate}
	%%
	\item All subjects,
	%%
	\item subdivision into \emph{male} and \emph{female},
	%%
	\item subdivision into \emph{age quantiles}, yielding the groups (21, 43.8] (43.8, 54] (54, 64] (64, 90],
	%%
	\item subdivision into \emph{size quantiles}, yielding the groups (139, 164], (164, 171], (171, 177], (177, 202].
	%%
\end{enumerate}
We plotted each group twice.
%%
The first plot shows all original variables.
%%
The second plot shows all categorical variables transformed into dichotomous dummy variables.
